Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2021-09-08 paper/paper.tex
add generic programming

\section{Introduction}
\blue{efficient counting maximal independent sets~\cite{Manne2013}}
In graph theory and combinatorics, there are many interesting but hard computational problems concerning properties of independent sets. An independent set is a set of vertices in a graph where no two vertices are adjacent to each other. More formally, given an undirected graph $G = (V,E)$, an independent set $I \subseteq V$ is a set that for any $u,v \in I$, there is no edge connecting $u$ and $v$ in $G$. The problem of finding an independent set of the largest possible size, the maximum independent set (MIS), is a paradigmatic NP-complete problem~\cite{Hastad1996}. The size of the MIS of a graph $G$ is called the independence number, denoted as $\alpha(G) \equiv \max_{I}|I|$.

Finding $\alpha(G)$ is a hard computational problem; moreover, it is NP-hard to even approximate $\alpha(G)$ within a factor $|V|^{1-\epsilon}$ for an arbitrarily small positive $\epsilon$. Naive exhaustive search for the MIS requires time $O(2^{|V|})$. More efficient exact algorithms include the branching algorithms and dynamic programming. Sophisticated branching algorithms can reduce the base of the exponential time scaling to, e.g., $1.1893^n n ^{O(1)}$~\cite{Xiao2017} \red{reading the abstract, I think the algorithm is $1.1996^{|V|} {|V|}^{O(1)}$. The $1.1893^{|V|} {|V|}^{O(1)}$ scaling is for graphs with maximum degree 6}. Dynamic programming approaches~\cite{Courcelle1990, Fomin2013} works better for graphs with a small treewidth $\text{tw}(G)$; these methods can produce algorithms of complexity $O(2^{\text{tw}(G)} \text{tw}(G) |V|)$.

The independent set problem is closely related to the clique problem and the vertex cover problem~\cite{Moore2011}; more concretely, an MIS of a graph $G$ corresponds to a maximum clique in the complement graph of $G$, and the complement vertex set of an MIS corresponds to a minimum vertex cover of the graph $G$. Together, these problems find a wide range of applications in scheduling, logistics, social network analysis, bioinformatics, wireless networks and telecommunication, map labelling, computer vision, etc.~\cite{Butenko2003Maximum, Wu2015review}. Besides the standard MIS problem, there are many other interesting problems pertaining to independent sets, such as the number of independent sets, independence polynomial, maximal independence polynomial, and enumeration of optimal and sub-optimal configurations. Some of these problems are of great interest in physics applications such as the hard-core lattice gas model~\cite{Dyre2016, Fernandes2007Monte} in statistical mechanics and the Rydberg hamiltonian with neutral atoms~\cite{Pichler2018} \red{cite experiment when it's ready}; they can, for example, be used to understand phase transitions and to identify harder graphs in an ensemble of graphs \red{cite experiment}.

@@

\section{Tensor networks}\red{I removed the superscript and subscript, since I think it's not necessary to introduce those.}
A tensor network~\cite{Cirac2020Matrix, Orus2014} can be viewed as a generalization of matrix multiplication to multi-dimensional tensor contraction. Let $A, B$ be two matrices; the matrix multiplication is defined as $C_{ik} = \sum_{j}A_{ij}B_{jk}$. Using the Einstein notation, we can write $C_{ik} = A_{ij}B_{jk}$, where $j$ is implicitly summed over. In the standard notation, each index can appear at most twice. A tensor network consists of multiple tensors performing the sum-product operation, i.e., contraction, and the tensor network can be represented as a multigraph with open edges by viewing a tensor as a vertex, a label pairing two tensors as an edge, and the remaining unpaired labels as open edges.

@@

\begin{example}
$C_{ijkl} = A_{jkm} B_{mil} V_{jm}$ is a tensor network representing $C_{ijkl} = \sum_{m}A_{jkm} B_{mil} V_{jm}$  \red{I've changed this. please check. The original was $C_{ijk} = A_{jkm} B_{mil} V_{jm}$ and $C_{ijk} = \sum_{ml}A_{jkm} B_{mil} V_{jm}$}. Its hypergraph representation is shown below, where we use different colors to annotate different hyperedges.

\centerline{\begin{tikzpicture}[
    dot/.style = {circle, fill, minimum size=#1,
                inner sep=0pt, outer sep=0pt},
    dot/.default = 6pt  % size of the circle diameter 
                    ]  
    \def\dx{0};
    \def\r{0.5cm}
    \def\ax{0}
    \def\ay{0}
    \def\bx{1}
    \def\by{1}
    \def\cx{1}
    \def\cy{-1}
    \node[color=white,fill=black,dot=\r] at (\ax+\dx,\ax) (a) {A};
    \node[color=white,fill=black,dot=\r] at (\bx+\dx,\by) (b) {B};
    \node[color=white,fill=black,dot=\r] at (\cx+\dx,\cy) (v) {V};
    \node at (\ax-0.8,\ay) (k) {k};
    \node at (\bx+0.4,\ay) (m) {m};
    \node at (\ax,\cy) (j) {j};
    \node at (\bx+1,\by) (l) {l};
    \node at (\bx-1,\by) (i) {i};
    \draw[color=blue] (i) -- (b);
    \draw[color=blue] (i) -- (\ax-1,\by);
    \draw[color=cyan,thick] (l) -- (b);
    \draw[color=violet,thick] (k) -- (a);
    \draw[color=violet,thick] (k) -- (\ax-1.5,\ay);
    \draw[color=black,thick] (b) -- (m);
    \draw[color=black,thick] (m) -- (a);
    \draw[color=black,thick] (m) -- (v);
    \draw[color=red,thick] (a) -- (j);
    \draw[color=red,thick] (v) -- (j);
    \draw[color=red,thick] (\ax-1,\cy) -- (j);
\end{tikzpicture}}
\end{example}

@@

In the spirit of generic programming, the elements of the tensors in the tensor network do not always need to be floating point numbers or integers. In fact, for the tensor network contraction to make sense, one only needs the tensor elements to form a commutative semiring. A semiring is a ring without additive inverse, while a commutative semiring is a semiring where the multiplication operation is commutative. To define a commutative semiring with the addition operation $\oplus$ and the multiplication operation $\odot$ on a set $R$, the following relations must hold for arbitrary three elements $a, b, c \in R$.
\begin{align*}
(a \oplus b) \oplus c = a \oplus (b \oplus c) & \hspace{5em}\text{$\triangleright$ commutative monoid $\oplus$ with identity $\mymathbb{0}$}\\
a \oplus \mymathbb{0} = \mymathbb{0} \oplus a = a &\\
a \oplus b = b \oplus a &\\
&\\
(a \odot b) \odot c = a \odot (b \odot c)  &   \hspace{5em}\text{$\triangleright$ commutative monoid $\odot$ with identity $\mymathbb{1}$}\\
a \odot  \mymathbb{1} =  \mymathbb{1} \odot a = a &\\
a \odot b = b \odot a &\\
&\\
a \odot (b\oplus c) = a\odot b \oplus a\odot c  &  \hspace{5em}\text{$\triangleright$ left and right distributive}\\
(a\oplus b) \odot c = a\odot c \oplus b\odot c &\\
&\\
a \odot \mymathbb{0} = \mymathbb{0} \odot a = \mymathbb{0}
\end{align*}
The requirement of being commutative is for the tensor contraction result to be independent of the contraction order, which could be relaxed in certain cases. In the following sections, we show how to compute the independence polynomial, the maximal independence polynomial, independence number, the number of independent sets, and enumeration of the MIS configurations of a general graph $G$ by designing tensor element types as commutative semirings while keeping the tensor network generic~\cite{Stepanov2014}. Table~\ref{tbl:generictypes} summarizes the problems that can be solved by various tensor element types.

\begin{table}[t!]\centering
\begin{minipage}{\columnwidth}
\ra{1.3}
        \begin{tabularx}{\textwidth}{bb}\toprule
            \hline
            \textbf{element type}     & \textbf{purpose} \\
            {real number}     & {counting all independent sets} \\
            {tropical number} (\Eq{eq:countingtropical})    & {finding the independence number} \\
            {tropical number with counting} (\Eq{eq:countingtropical})     & {finding the independence number and the number of MISs} \\
            {tropical number with configurations} (\Eq{eq:singleconfig})     & {finding the independence number and one MIS} \\
            {tropical number with sets} (\Eq{eq:set})    & {finding the independence number and enumeration of all MISs} \\
            {polynomial} (\Eq{eq:polynomial})     & {computing the independence polynomial exactly} \\
            {truncated polynomial} (\Eq{eq:max2poly})     & {counting MISs and independent sets of size $\alpha(G)-1$} \\
            {complex number}     & {fitting the independence polynomial with fast Fourier transform} \\
            {finite-field algebra (\Eq{eq:finitefield})} & {fitting the independence polynomial exactly using number theory} \\
            \bottomrule
        \end{tabularx}
    \caption{Tensor element types and their purposes in calculating various independent set properties.}\label{tbl:generictypes}
\end{minipage}
\end{table}

\section{Independence polynomial and maximal independence polynomial}
\subsection{Independence polynomial}\label{sec:indpoly}
The independence polynomial is an important graph polynomial that contains the counting information of independent sets. It is defined as
\begin{equation}
I(G, x) = \sum_{k=0}^{\alpha(G)} a_k x^k,
\end{equation}
where $a_k$ is the number of independent sets of size $k$ in $G$, and $\alpha(G)$ is the independence number. The total number of independent sets is thus equal to $I(G, 1)$. The problem of computing the independence polynomial belongs to the complexity class \#P-hard. One approach to exactly compute the independence polynomial requires a computational time of $O(1.442^{|V|})$~\cite{Ferrin2014}\blue{I am not sure about this complexity, this is based on the naive analysis of theorem 2.2 in ~\cite{Ferrin2014}}. There are some interesting works on efficiently approximating the independence polynomial~\cite{Harvey2018}, but in this work, we focus on exact methods to compute the independence polynomial.

The independence polynomial is closely related to the matching polynomial, the clique polynomial, and the vertex cover polynomial. In fact, the independence polynomial can be viewed as a generalization of the matching polynomial, since the matching polynomial of a graph $G$ is the same as the independence polynomial of the line graph of $G$~\cite{levit2005Independence}. Thus, our algorithm to compute the independence polynomial can also be used to compute the above graph polynomials. Some properties of the independence polynomial, such as the uni-modality, log-concavity, and roots of the polynomial, are well studied in the literature for special classes of graphs~\cite{levit2005Independence}. We hope our algorithms to calculate the independence polynomial can also help further research in these directions.

To compute the independence polynomial of a graph $G$, we encode it to a tensor network. On the vertex $i$ of the graph $G$, we place a rank-one vertex tensor, $W(x_i)$, of size $2$ parametrized by $x_{i}$ and a rank-two edge tensor, $B$, of size $2 \times 2$ on an edge $(i,j)$
\red{I removed the indices $s_{i}$ etc, on the right of the equations, since that looks confusing to me. If I understand correctly, $s_{i}, s_{j}$ just denotes the index of the vector/matrix.}
\begin{equation}
    W(x_i)_{s_i} = \left(\begin{matrix}
        1 \\
        x_i
    \end{matrix}\right),
    \qquad \quad 
       B_{s_i s_j} = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right), \label{eq:tensor}
\end{equation}
where the tensor index $s_i$ is a boolean variable; $s_{i} = 1$ corresponds that the vertex $i$ is in the independent set, and $s_{i} = 0$ means otherwise. Let us assume the tensor elements are real numbers for now. The edge tensor element $B_{s_{i}=1, s_{j}=1} = 0$ encodes the independent set constraint, meaning vertex $i$ and $j$ cannot be both in the independent set if they are connected by an edge $(i,j)$. The contraction of the tensor network representing $G$ gives
\begin{equation}
    P(G, \{x_1, x_{2}, \ldots,x_{|V|}\}) = \sum\limits_{s_1, s_2, \ldots, s_{|V|} = 0}^{1} \prod\limits_{i=1}^{|V|} W(x_i)_{s_i} \prod\limits_{(i,j) \in E(G)} B_{s_i s_j},
\end{equation}
where the summation runs over all vertex configurations $\{s_1, s_{2}, \ldots,s_{|V|}\}$ and accumulates the product of tensor elements to the scalar output $P$ (see Example \ref{eg:tensorcontraction} for a concrete example). In the special case of $x_i = x$, the contraction result directly corresponds to the independence polynomial. The connection can be understood as follows: the product over vertex tensor elements produces a factor $x^k$, where $k=\sum_i s_i$ counts the set size, and the product over edge tensor elements gives a factor $1$ for a configuration being in an independent set and $0$ otherwise. In the implementation of the algorithm, we benefit from recent advances in tensor network contraction, where researchers working on quantum circuit simulation evaluate the tensor network by pairwise contracting tensors in a good heuristic order~\cite{Gray2021, Pan2021}. A good contraction order can reduce the time complexity significantly, at the cost of having a space overhead of $O(2^{\text{tw}(G)})$~\cite{Markov2008}. The pairwise tensor contraction also makes it possible to utilize basic linear algebra subprograms (BLAS) functions to speed up the computation for certain tensor element types.

@@

\subsection{Maximal independence polynomial} \red{It might be helpful to add a figure of the tensor network for computing the maximal independence polynomial.} Instead of counting all independent sets, the maximal independence polynomial counts the number of maximal independent sets of various sizes~\cite{Hu2017On}. Concretely, it is defined as
\begin{equation}
I_{\rm max}(G, x) = \sum_{k=0}^{\alpha(G)} b_k x^k,
\end{equation}
where $b_k$ is the number of maximal independent sets of size $k$ in $G$. Obviously, $b_{k} \leq a_{k}$ and $b_{\alpha(G)} = a_{\alpha(G)}$. $I_{\rm max}(G, 1)$ counts the total number of maximal independent sets~\cite{Gaspers2012On, Manne2013}, where the fastest algorithm currently has a runtime of $O(1.3642^{|V|})$~\cite{Gaspers2012On}. For the problem of finding the MIS, $b_{k}$ counts the number of local optimum at size $k < \alpha(G)$, and can, in some cases, provide hints on the difficulty of finding the MIS using local algorithms~\red{cite experiment}. The uni-modality, log-concavity, and real-rootness properties of the maximal independence polynomial for special classes of graphs have also been studied~\cite{Hu2017On}.

@@

With the tensors defined as $T(x_v)$, we can perform a similar computation of contracting the tensor network with the same tensor element types as described in the previous section~\ref{sec:indpoly}, the result of which then produces the maximal independence polynomial. The computational complexity of this new tensor network contraction is often greater than the one for computing the independence polynomial. However, in most sparse graphs, this tensor network contraction approach is still significantly faster than enumerating all the maximal cliques on its complement graph using the Bron-Kerbosch algorithm~\cite{Bron1973Algorithm}, which is the standard algorithm that we are aware of to compute the maximal independence polynomial.

@@

\section{Enumeration of configurations and enumeration bounds}
\subsection{Enumeration of MIS configurations}
The enumeration problems of independent sets are also interesting and has been studied extensively in the literature~\cite{Bron1973Algorithm, Eppstein2010Listing, Johnson1988On}, including, for example, the enumeration of all independent sets, the enumeration of all maximal independent sets, or the enumeration of all MISs. The standard algorithm to enumerate all maximal independent sets is the Bron-Kerbosch algorithm~\cite{Bron1973Algorithm}, which, of course, includes the enumeration of all MISs and can be used to enumerate all independent sets as well. Here, we adapt the tensor element types for different enumeration problems. For example, our tensor-network based algorithm to enumerate all MISs is expectedly much faster and more space-efficient than the Bron-Kerbosch algorithm, which lists all maximal independent sets.

@@

where $s$ and $t$ are each a set of $|V|$-bit strings and $\lor^\circ$ is the bitwise OR operation over two bit strings. \red{This is the bitwise OR operation right? I can't seem to find online what you call Hadamard. Is there a definition somewhere?} \red{it would be useful to give an example of the operations above, $s = ... , t = ..., s \oplus t = ..., s \odot t = ...$.} \red{do we need to define $ \{\} \lor^\circ \sigma  = \sigma \lor^\circ \{\} = \{\}$?} The variable $x_{i}$ in the vertex tensor is initialized to $x_i = \{\boldsymbol{e}_{i}\}$, where $\boldsymbol{e}_i$ is the standard basis vector of size $|V|$ and having 1 at index $i$. The vertex and edge tensors are thus

@@

Contraction of the tensor network enumerates all the independent sets. To enumerate only the MISs, we can combine the above algebra \Eq{eq:set} with the tropical algebra in \Eq{eq:countingtropical} and define $s(k) = (s_{k}, k)$, where the first element follows the algebra in \Eq{eq:set} and the second element follows the max-plus tropical algebra. The combined operations become: 

@@

    \mymathbb{1} &= (\{0^{\otimes |V|}\}, 0). \label{eq:countingtropical}

@@

The contraction of the corresponding tensor network yields an enumeration of all MIS configurations. If one is interested in obtaining only a single MIS configuration, one can just keep a single configuration in the intermediate computations to save the computational effort. Here is a new algebra defined on the bit strings, replacing the sets of bit strings in \Eq{eq:set}, 

@@

\section{Why not introducing $\delta$ tensors}\label{app:tensorbad}

>>>>

\section{Introduction}
\blue{overlap gap property~\cite{Gamarnik2013, Gamarnik2019}, independence polynomial at -1~\cite{Bousquet2008,Levit2009}, treewidth of 3-regular graphs ~\cite{Fomin2006},
branching algorithm~\cite{Tarjan1977,Robson1986}
}
In graph theory and combinatorics, there are many interesting but hard computational problems concerning properties of independent sets.
An independent set is a set of vertices in a graph where no two vertices are adjacent to each other.
More formally, given an undirected graph $G = (V,E)$, an independent set $I \subseteq V$ is a set that for any $u,v \in I$, there is no edge connecting $u$ and $v$ in $G$.
The problem of finding an independent set of the largest possible size, the maximum independent set (MIS), is a paradigmatic NP-complete problem~\cite{Hastad1996}.
The size of the MIS of a graph $G$ is called the independence number, denoted as $\alpha(G) \equiv \max_{I}|I|$.

Finding $\alpha(G)$ is a hard computational problem; moreover, it is NP-hard to even approximate $\alpha(G)$ within a factor $|V|^{1-\epsilon}$ for an arbitrarily small positive $\epsilon$.
Naive exhaustive search for the MIS requires time $O(2^{|V|})$. More efficient exact algorithms include the branching algorithms and dynamic programming.
Sophisticated branching algorithms can reduce the base of the exponential time scaling to, e.g., $1.1996^n n ^{O(1)}$~\cite{Xiao2017}.
Dynamic programming approaches~\cite{Courcelle1990, Fomin2013} works better for graphs with a small treewidth $\text{tw}(G)$;
these methods can produce algorithms of complexity $O(2^{\text{tw}(G)} \text{tw}(G) |V|)$.

The independent set problem is closely related to the clique problem and the vertex cover problem~\cite{Moore2011};
more concretely, an MIS of a graph $G$ corresponds to a maximum clique in the complement graph of $G$,
and the complement vertex set of an MIS corresponds to a minimum vertex cover of the graph $G$.
Together, these problems find a wide range of applications in scheduling, logistics, social network analysis, bioinformatics, wireless networks and telecommunication,
map labelling, computer vision, etc.~\cite{Butenko2003, Wu2015}.
Besides the standard MIS problem, there are many other interesting problems pertaining to independent sets, such as the number of independent sets,
independence polynomial, maximal independence polynomial, and enumeration of optimal and sub-optimal configurations.
Some of these problems are of great interest in physics applications such as the hard-core lattice gas model~\cite{Dyre2016, Fernandes2007}
in statistical mechanics and the Rydberg hamiltonian with neutral atoms~\cite{Pichler2018} \red{cite experiment when it's ready};
they can, for example, be used to understand phase transitions and to identify harder graphs in an ensemble of graphs \red{cite experiment}.

@@

\section{Tensor networks}
A tensor network~\cite{Cirac2020, Orus2014} can be viewed as a generalization of matrix multiplication to multi-dimensional tensor contraction.
Let $A, B$ be two matrices; the matrix multiplication is defined as $C_{ik} = \sum_{j}A_{ij}B_{jk}$.
Using the Einstein notation, we can write $C_{ik} = A_{ij}B_{jk}$, where $j$ is implicitly summed over.
In the standard notation, each index can appear at most twice. A tensor network consists of multiple tensors performing the sum-product operation,
i.e., contraction, and the tensor network can be represented as a multigraph with open edges by viewing a tensor as a vertex,
a label pairing two tensors as an edge, and the remaining unpaired labels as open edges.

@@

\begin{example}
$C_{ijk} = A_{jkm} B_{mil} V_{jm}$ is a tensor network representing $C_{ijk} = \sum_{ml}A_{jkm} B_{mil} V_{jm}$ 
\red{I've changed this. please check. The original was $C_{ijk} = A_{jkm} B_{mil} V_{jm}$ and $C_{ijk} = \sum_{ml}A_{jkm} B_{mil} V_{jm}$}.
\blue{changed back because $l$ is not an open edge. Is it still confusing if I introduce open circles for open edges?}
Its hypergraph representation is shown below, where we use different colors to annotate different hyperedges and open circles to denote open edges.

\centerline{\begin{tikzpicture}[
    dot/.style = {circle, fill, minimum size=#1,
                inner sep=0pt, outer sep=0pt},
    dot/.default = 6pt  % size of the circle diameter 
                    ]  
    \def\dx{0};
    \def\r{0.5cm}
    \def\sr{0.15cm}
    \def\ax{0}
    \def\ay{0}
    \def\bx{1}
    \def\by{1}
    \def\cx{1}
    \def\cy{-1}
    \node[color=white,fill=black,dot=\r] at (\ax+\dx,\ax) (a) {A};
    \node[color=white,fill=black,dot=\r] at (\bx+\dx,\by) (b) {B};
    \node[color=white,fill=black,dot=\r] at (\cx+\dx,\cy) (v) {V};
    \node[color=white,draw=black,dot=\sr] at (\ax-1,\by) (o1) {};
    \node[color=white,draw=black,dot=\sr] at (\ax-1.5,\ay) (o2) {};
    \node[color=white,draw=black,dot=\sr] at (\ax-1,\cy) (o3) {};
    \node at (\ax-0.8,\ay) (k) {k};
    \node at (\bx+0.4,\ay) (m) {m};
    \node at (\ax,\cy) (j) {j};
    \node at (\bx+1,\by) (l) {l};
    \node at (\bx-1,\by) (i) {i};
    \draw[color=blue,thick] (i) -- (b);
    \draw[color=blue,thick] (i) -- (o1);
    \draw[color=cyan,thick] (l) -- (b);
    \draw[color=violet,thick] (k) -- (a);
    \draw[color=violet,thick] (k) -- (o2);
    \draw[color=black,thick] (b) -- (m);
    \draw[color=black,thick] (m) -- (a);
    \draw[color=black,thick] (m) -- (v);
    \draw[color=red,thick] (a) -- (j);
    \draw[color=red,thick] (v) -- (j);
    \draw[color=red,thick] (o3) -- (j);
\end{tikzpicture}}
\end{example}

@@

In previous works combining tensor network and combinatoric problems~\cite{},
the elements in tensor networks are limited to regular number types such as floating point numbers and integers.
As the development of modern compiling technology, we do not need to limit our imaginations to regular number types anymore.
The key technology is called the generic programming:

\begin{definition}[Generic programming]
   Generic programming is an approach to programming that focuses on designing algorithms and data structures so that they work in the most general setting without loss of efficiency. ~\cite{Stepanov2014}
\end{definition}

This definition of generic programming contains two major pieces, one is a single program for the general setting, and another is performance.
To understand the first point, suppose we want to write a function to take the power of an element $f(x, n) := x^n$,
the generic programming approach does not assume $x$ being a regular number,
instead it treats $x$ as an element with an associative multiplication operation $\odot$ and multiplicative identity $\mymathbb{1}$.
In such a way, when the program takes a matrix as input, it computes the matrix power without rewriting this function for matrices.
The second point about performance basically rules out dynamic typed languages like Python,
because in these languages, the efficiency of a generic type is not guaranteed.
Imagine the speed of computing the matrix multiplication between two numpy arrays with python objects as elements.
The generic programming that we want to discuss is those in static types languages like C++, and Julia~\cite{Bezanson2012}.
C++ uses templates for generic programming while Julia has just in time compiling and multiple dispatch.
When the program ``sees'' a new input type, the compiler recompiles the generic program for the new type.
A lot optimization can be done during the compiling, like inlining immutable elements with fixed sizes in an array.
In Julia, these inlined arrays can even be compiled to GPU devices for faster computation~\cite{Besard2018}.

This leads us to think about what is the most general element type that allowed in a tensor network contraction program.
We find as long as the tensor elements form a commutative semiring, the tensor network contraction result can be contraction order independent.
A semiring is a ring without additive inverse, while a commutative semiring is a semiring where the multiplication operation is commutative.
To define a commutative semiring with the addition operation $\oplus$ and the multiplication operation $\odot$ on a set $R$, the following relations must hold for arbitrary three elements $a, b, c \in R$.
\begin{align*}
(a \oplus b) \oplus c = a \oplus (b \oplus c) & \hspace{5em}\text{$\triangleright$ commutative monoid $\oplus$ with identity $\mymathbb{0}$}\\
a \oplus \mymathbb{0} = \mymathbb{0} \oplus a = a &\\
a \oplus b = b \oplus a &\\
&\\
(a \odot b) \odot c = a \odot (b \odot c)  &   \hspace{5em}\text{$\triangleright$ commutative monoid $\odot$ with identity $\mymathbb{1}$}\\
a \odot  \mymathbb{1} =  \mymathbb{1} \odot a = a &\\
a \odot b = b \odot a &\\
&\\
a \odot (b\oplus c) = a\odot b \oplus a\odot c  &  \hspace{5em}\text{$\triangleright$ left and right distributive}\\
(a\oplus b) \odot c = a\odot c \oplus b\odot c &\\
&\\
a \odot \mymathbb{0} = \mymathbb{0} \odot a = \mymathbb{0}
\end{align*}
The requirement of being commutative is for the tensor contraction result to be independent of the contraction order, which could be relaxed in certain cases. In the following sections, we show how to compute the independence polynomial,
the maximal independence polynomial, independence number, the number of independent sets, and enumeration of the MIS configurations of a general graph $G$ by designing tensor element types as commutative semirings while keeping the tensor network generic~\cite{Stepanov2014}.
Table~\ref{tbl:generictypes} summarizes the problems that can be solved by various tensor element types.

\begin{table}[t!]\centering
\begin{minipage}{\columnwidth}
\ra{1.3}
        \begin{tabularx}{\textwidth}{bb}\toprule
            \hline
            \textbf{element type}     & \textbf{purpose} \\
            {real number}     & {counting all independent sets} \\
            {tropical number} (\Eq{eq:countingtropical})    & {finding the independence number} \\
            {tropical number with counting} (\Eq{eq:countingtropical})     & {finding the independence number and the number of MISs} \\
            {tropical number with configurations} (\Eq{eq:singleconfig})     & {finding the independence number and one MIS} \\
            {tropical number with sets} (\Eq{eq:countingtropicalset})    & {finding the independence number and enumeration of all MISs} \\
            {polynomial} (\Eq{eq:polynomial})     & {computing the independence polynomial exactly} \\
            {truncated polynomial} (\Eq{eq:max2poly})     & {counting MISs and independent sets of size $\alpha(G)-1$} \\
            {complex number}     & {fitting the independence polynomial with fast Fourier transform} \\
            {finite-field algebra (\Eq{eq:finitefield})} & {fitting the independence polynomial exactly using number theory} \\
            \bottomrule
        \end{tabularx}
    \caption{Tensor element types and their purposes in calculating various independent set properties.}\label{tbl:generictypes}
\end{minipage}
\end{table}

\section{Independence polynomial and maximal independence polynomial}
\subsection{Independence polynomial}\label{sec:indpoly}
The independence polynomial is an important graph polynomial that contains the counting information of independent sets. It is defined as
\begin{equation}
I(G, x) = \sum_{k=0}^{\alpha(G)} a_k x^k,
\end{equation}
where $a_k$ is the number of independent sets of size $k$ in $G$, and $\alpha(G)$ is the independence number. The total number of independent sets is thus equal to $I(G, 1)$.
The problem of computing the independence polynomial belongs to the complexity class \#P-hard. One approach to exactly compute the independence polynomial requires a computational time of $O(1.442^{|V|})$~\cite{Ferrin2014}
\blue{I am not sure about this complexity, this is based on the naive analysis of theorem 2.2 in ~\cite{Ferrin2014}}. There are some interesting works on efficiently approximating the independence polynomial~\cite{Harvey2018}, but in this work, we focus on exact methods to compute the independence polynomial.

The independence polynomial is closely related to the matching polynomial, the clique polynomial, and the vertex cover polynomial.
In fact, the independence polynomial can be viewed as a generalization of the matching polynomial, since the matching polynomial of a graph $G$ is the same as the independence polynomial of the line graph of $G$~\cite{Levit2005}.
Thus, our algorithm to compute the independence polynomial can also be used to compute the above graph polynomials. Some properties of the independence polynomial, such as the uni-modality, log-concavity, and roots of the polynomial, are well studied in the literature for special classes of graphs~\cite{Levit2005}.
We hope our algorithms to calculate the independence polynomial can also help further research in these directions.

To compute the independence polynomial of a graph $G$, we encode it to a tensor network. On the vertex $i$ of the graph $G$, we place a rank-one vertex tensor, $W(x_i)$, of size $2$ parametrized by $x_{i}$ and a rank-two edge tensor, $B$, of size $2 \times 2$ on an edge $(i,j)$
\red{I removed the indices $s_{i}$ etc, on the right of the equations, since that looks confusing to me. If I understand correctly, $s_{i}, s_{j}$ just denotes the index of the vector/matrix.}
\blue{To be consistent, we should also remove the index on the LHS. It is better not removing the label information.}
\begin{equation}
    W(x_i)_{s_i} = \left(\begin{matrix}
        1 \\
        x_i
    \end{matrix}\right),
    \qquad \quad 
       B_{s_i s_j} = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right), \label{eq:tensor}
\end{equation}
where the tensor index $s_i$ is a boolean variable; $s_{i} = 1$ corresponds that the vertex $i$ is in the independent set, and $s_{i} = 0$ means otherwise. Let us assume the tensor elements are real numbers for now.
The edge tensor element $B_{s_{i}=1, s_{j}=1} = 0$ encodes the independent set constraint, meaning vertex $i$ and $j$ cannot be both in the independent set if they are connected by an edge $(i,j)$. The contraction of the tensor network representing $G$ gives
\begin{equation}
    P(G, \{x_1, x_{2}, \ldots,x_{|V|}\}) = \sum\limits_{s_1, s_2, \ldots, s_{|V|} = 0}^{1} \prod\limits_{i=1}^{|V|} W(x_i)_{s_i} \prod\limits_{(i,j) \in E(G)} B_{s_i s_j},
\end{equation}
where the summation runs over all vertex configurations $\{s_1, s_{2}, \ldots,s_{|V|}\}$ and accumulates the product of tensor elements to the scalar output $P$ (see Example \ref{eg:tensorcontraction} for a concrete example).
In the special case of $x_i = x$, the contraction result directly corresponds to the independence polynomial.
The connection can be understood as follows: the product over vertex tensor elements produces a factor $x^k$, where $k=\sum_i s_i$ counts the set size,
and the product over edge tensor elements gives a factor $1$ for a configuration being in an independent set and $0$ otherwise. In the implementation of the algorithm, we benefit from recent advances in tensor network contraction,
where researchers working on quantum circuit simulation evaluate the tensor network by pairwise contracting tensors in a good heuristic order~\cite{Gray2021, Pan2021}.
A good contraction order can reduce the time complexity significantly, at the cost of having a space overhead of $O(2^{\text{tw}(G)})$~\cite{Markov2008}.
The pairwise tensor contraction also makes it possible to utilize basic linear algebra subprograms (BLAS) functions to speed up the computation for certain tensor element types.

@@

\subsection{Maximal independence polynomial}
\red{It might be helpful to add a figure of the tensor network for computing the maximal independence polynomial.} 
Sometimes people are interested in knowing maximal solutions to understand why their programs are trapped in a local minimal.
In this paper, the word ``maximal'' is different with the ``maximum'' discussed the previous discussion in such a way that it is not necessarily refer to the gloal optimum.
Instead of counting all independent sets, the maximal independence polynomial counts the number of maximal independent sets of various sizes~\cite{Hu2017}.
 Concretely, it is defined as
\begin{equation}
I_{\rm max}(G, x) = \sum_{k=0}^{\alpha(G)} b_k x^k,
\end{equation}
where $b_k$ is the number of maximal independent sets of size $k$ in $G$.
Obviously, $b_{k} \leq a_{k}$ and $b_{\alpha(G)} = a_{\alpha(G)}$. $I_{\rm max}(G, 1)$ counts the total number of maximal independent sets~\cite{Gaspers2012, Manne2013}, where the fastest algorithm currently has a runtime of $O(1.3642^{|V|})$~\cite{Gaspers2012}.
For the problem of finding the MIS, $b_{k}$ counts the number of local optimum at size $k < \alpha(G)$, and can, in some cases, provide hints on the difficulty of finding the MIS using local algorithms~\red{cite experiment}.
The uni-modality, log-concavity, and real-rootness properties of the maximal independence polynomial for special classes of graphs have also been studied~\cite{Hu2017}.

@@

With the tensors defined as $T(x_v)$, we can perform a similar computation of contracting the tensor network with the same tensor element types as described in the previous section~\ref{sec:indpoly}, the result of which then produces the maximal independence polynomial.
The computational complexity of this new tensor network contraction is often greater than the one for computing the independence polynomial.
However, in most sparse graphs, this tensor network contraction approach is still significantly faster than enumerating all the maximal cliques on its complement graph using the Bron-Kerbosch algorithm~\cite{Bron1973}, which is the standard algorithm that we are aware of to compute the maximal independence polynomial.

@@

\section{Enumeration of configurations and enumeration bounds}
\subsection{Enumeration of MIS configurations}
The enumeration problems of independent sets are also interesting and has been studied extensively in the literature~\cite{Bron1973, Eppstein2010, Johnson1988}, including,
for example, the enumeration of all independent sets, the enumeration of all maximal independent sets, or the enumeration of all MISs.
The standard algorithm to enumerate all maximal independent sets is the Bron-Kerbosch algorithm~\cite{Bron1973}, which, of course, includes the enumeration of all MISs and can be used to enumerate all independent sets as well.
Here, we adapt the tensor element types for different enumeration problems.
For example, our tensor-network based algorithm to enumerate all MISs is expectedly much faster and more space-efficient than the Bron-Kerbosch algorithm, which lists all maximal independent sets.

@@

where $s$ and $t$ are each a set of $|V|$-bit strings and $\lor^\circ$ is the bitwise OR operation over two bit strings. \red{This is the bitwise OR operation right? I can't seem to find online what you call Hadamard. Is there a definition somewhere?}
\red{it would be useful to give an example of the operations above, $s = ... , t = ..., s \oplus t = ..., s \odot t = ...$.} \red{do we need to define $ \{\} \lor^\circ \sigma  = \sigma \lor^\circ \{\} = \{\}$?}
The variable $x_{i}$ in the vertex tensor is initialized to $x_i = \{\boldsymbol{e}_{i}\}$, where $\boldsymbol{e}_i$ is the standard basis vector of size $|V|$ and having 1 at index $i$. The vertex and edge tensors are thus

@@

Contraction of the tensor network enumerates all the independent sets.
To enumerate only the MISs, we can combine the above algebra \Eq{eq:set} with the tropical algebra in \Eq{eq:countingtropical} and define $s(k) = (s_{k}, k)$, where the first element follows the algebra in \Eq{eq:set} and the second element follows the max-plus tropical algebra.
The combined operations become: 

@@

    \mymathbb{1} &= (\{0^{\otimes |V|}\}, 0). \label{eq:countingtropicalset}

@@

The contraction of the corresponding tensor network yields an enumeration of all MIS configurations.
If one is interested in obtaining only a single MIS configuration, one can just keep a single configuration in the intermediate computations to save the computational effort.
Here is a new algebra defined on the bit strings, replacing the sets of bit strings in \Eq{eq:set}, 

@@

\section{Why not introducing \texorpdfstring{$\delta$} tensors}\label{app:tensorbad}
